---

title: "R Notebook"
output:
  html_document:
    df_print: paged
  html_notebook: default

---

The hypothesis we are testing is whether VR treatment had any effect on GAD scores:

Null hypothesis is equal to 0
Alternative hypothesis is not equal to 0

---

We started by cleaning the data first checking the relevant data are numeric


```{r}
# Check if pre-trial values numeric
is.numeric(GROUP_3_2023_GCA_RESULTS_CONTROL.csv$pretrial_GAD)

```


```{r}
# Check if control post-trial data is numeric
Ctrposttrial_GAD <- c(GROUP_3_2023_GCA_RESULTS_CONTROL.csv$posttrial_GAD)
is.numeric(Ctrposttrial_GAD)

```


```{r}
# Check if experiment pre-trial data is numeric
dataCA1Experiment <- read.csv("GROUP_3_2023_GCA_RESULTS_EXPERIMENTAL.csv")
Exppretrial_GAD <- c(dataCA1Experiment$pretrial_GAD)
is.numeric(Exppretrial_GAD)

```


```{r}
# Check experiment post-trial data is numeric
dataCA1Experiment <- read.csv("GROUP_3_2023_GCA_RESULTS_EXPERIMENTAL.csv")
Expposttrial_GAD <- c(dataCA1Experiment$posttrial_GAD)
is.numeric(Expposttrial_GAD)

```

There were outliers in the pre-trial control data and the post-trial experiment data with negative values.

```{r}
# Boxplot showing outlier in control pre-trial GAD scores
boxplot(GROUP_3_2023_GCA_RESULTS_CONTROL.csv$pretrial_GAD)
```

```{r}
# Boxplot of experiment post-trial data showing outlier
boxplot(Expposttrial_GAD)
```


We cleaned the data and examined histograms of the positive date to confirm this was normally distributed

```{r}
# Get positive values
CtrPos_pretrial_GAD <- pretrial_GAD[pretrial_GAD >= 0]
CtrPos_pretrial_GAD

```
The histograms show data once cleaned is normally distributed and unimodal. 

```{r}
# Histogram of positive pretrial GAD values
pretrial_GAD <- c(GROUP_3_2023_GCA_RESULTS_CONTROL.csv$pretrial_GAD)
pretrial_GAD
CtrPos_pretrial_GAD <- pretrial_GAD[pretrial_GAD >= 0]
CtrPos_pretrial_GAD
hist(CtrPos_pretrial_GAD,
     main = "Control positive pre-trial GAD scores",
     xlab = "Positive pre-trial GAD scores")


```

```{r}
# Get positive experiment post-trial values
ExpPos_posttrial_GAD <- Expposttrial_GAD[Expposttrial_GAD >= 0]
ExpPos_posttrial_GAD
is.numeric(ExpPos_posttrial_GAD)
```

```{r}
# Histogram of postive experiment post trial values
hist(ExpPos_posttrial_GAD,
main = "Experiment positive post-trial GAD scores",
xlab = "Positive post-trial GAD scores")

```

Because the sample sizes were relatively small, we also ran Anderson Darling tests for normality. These produced unexpected results, possibly because the test can be affected by ties in the data. This needs further investigation. 

```{r}
# Anderson Darling test for normality - Positive pretrial GAD values for control data

#install and load nortest library
install.packages('nortest')
library(nortest)

ad.test(CtrPos_pretrial_GAD)
```

Anderson-Darling normality test for positive pre-trial GAD values for control data

data:  CtrPos_pretrial_GAD
A = 1.0512, p-value = 0.008746

A p value below .05 leads us to reject the null hypothesis that the data are normally distributed. This seems to contradict the appearance of a normal distribution shown in the histogram



```{r}
# Anderson Darling test for normality - Post-trial GAD values for control data

#load nortest library
library(nortest)

ad.test(Ctrposttrial_GAD)
```
A p value greater than .05 indicates the post-trial GAD values for control data are normally distributed


```{r}
# Anderson Darling test for normality - Pre-trial GAD values for experiment data

#load nortest library
library(nortest)

ad.test(Exppretrial_GAD)
```
P value less than .05 indicates experiment pre-trial data are not normally distributed

```{r}
# Anderson Darling test for normality - Positive Post-trial GAD values for experiment data

#load nortest library
library(nortest)

ad.test(ExpPos_posttrial_GAD)
```
A p value less that .05 indicates positive Post-trial GAD values for experiment data are not normally distributed

We decided to proceed with the t-tests as the histograms clearly show the positive data is normally distributed.


```{r}
# Two Sample t-test
t.test(Ctrposttrial_GAD, ExpPos_posttrial_GAD,var.equal = TRUE)

```
The p value is greater than .05 which indicates we should accept the null hypothesis that VR treatment had no effect on GAD scores.




