---

title: "R Notebook"
output:
  html_document:
    df_print: paged
  html_notebook: default

---
Contents

Abstract .............
Introduction .........
Method ...............
Results ..............
Discussion ...........
References ...........

---
Abstract

Aim and Rationale
The Aim of this study is to research a possibility of a more effective way to treat anxiety 
disorder using virtual headsets and environments.

Participants and setting
The study consisted of 180 participants divided into control and experimental groups. The participants were randomly sampled from the clinic client list and randomly assigned to the 2 groups. The participants were all ages between 18 - 40 years.

Experiment design
The 2 groups underwent 12 weeks of treatment for 50 minutes per week with a therapist. 
During that time the participants underwent either traditional CBT or used the VR app.

Results gathering
The results were gathered using the self-report mechanism at the beginning and end of the study.

Findings
This report finds that the null hypothesis should be accepted as the VR app had no evident effect on Generalized Anxiety Disorder 7-item scale scores

---
Introduction

As of 2019, according to the World Health Organization, approximately 301 million people worldwide suffer from anxiety disorders. Anxiety disorders are a group of mental disorders characterized by feelings of anxiety and fear. These feelings may cause physical symptoms such as a racing heart and shakiness. Anxiety is a natural response to stress, but for individuals with an anxiety disorder, the fear or nervousness does not go away and can get worse over time, affecting daily activities such as job performance, school-work, and relationships.

Triggers for anxiety disorders can range from environmental factors, genetics, brain chemistry, medical factors, to a combination of these. There are several types of anxiety disorders, including generalized anxiety disorder, panic disorder, specific phobias, agoraphobia, social anxiety disorder, and separation anxiety disorder. Treatment options for these disorders include a combination of psychotherapy, behavioral therapy,and medications.

The report below will test the null hypopthesis that the VR treatment had no effect on Generalized Anxiety Disorder 7-item scale scores

---
Method


We started by cleaning the data first checking the relevant data are numeric


```{r}
# Check if pre-trial values numeric
is.numeric(GROUP_3_2023_GCA_RESULTS_CONTROL.csv$pretrial_GAD)

```


```{r}
# Check if control post-trial data is numeric
Ctrposttrial_GAD <- c(GROUP_3_2023_GCA_RESULTS_CONTROL.csv$posttrial_GAD)
is.numeric(Ctrposttrial_GAD)

```


```{r}
# Check if experiment pre-trial data is numeric
dataCA1Experiment <- read.csv("GROUP_3_2023_GCA_RESULTS_EXPERIMENTAL.csv")
Exppretrial_GAD <- c(dataCA1Experiment$pretrial_GAD)
is.numeric(Exppretrial_GAD)

```


```{r}
# Check experiment post-trial data is numeric
dataCA1Experiment <- read.csv("GROUP_3_2023_GCA_RESULTS_EXPERIMENTAL.csv")
Expposttrial_GAD <- c(dataCA1Experiment$posttrial_GAD)
is.numeric(Expposttrial_GAD)

```

There were outliers in the pre-trial control data and the post-trial experiment data with negative values.

```{r}
# Boxplot showing outlier in control pre-trial GAD scores
boxplot(GROUP_3_2023_GCA_RESULTS_CONTROL.csv$pretrial_GAD)

```

```{r}
# Boxplot of experiment post-trial data showing outlier
boxplot(Expposttrial_GAD)
```


We cleaned the data and examined histograms of the positive data to confirm this was normally distributed

```{r}
# Get positive values
CtrPos_pretrial_GAD <- pretrial_GAD[pretrial_GAD >= 0]
CtrPos_pretrial_GAD

```
The histograms show data once cleaned is normally distributed and unimodal. 

```{r}
# Histogram of post-trial GAD values for control group
Ctrposttrial_GAD <- c(GROUP_3_2023_GCA_RESULTS_CONTROL.csv$posttrial_GAD)
Ctrposttrial_GAD
hist(Ctrposttrial_GAD,
     main = "Control Post-trial GAD scores",
     xlab = "Post-trial GAD scores")


```

```{r}
# Get positive experiment post-trial values
ExpPos_posttrial_GAD <- Expposttrial_GAD[Expposttrial_GAD >= 0]
is.numeric(ExpPos_posttrial_GAD)
```

```{r}
# Histogram of postive experiment post trial values
hist(ExpPos_posttrial_GAD,
main = "Experiment positive post-trial GAD scores",
xlab = "Positive post-trial GAD scores")

```

Because the sample sizes were relatively small, we also ran Anderson Darling tests for normality on the post-trial data in advance of carrying out t-tests.


```{r}
# Anderson Darling test for normality - Post-trial GAD values for control data

#load nortest library
library(nortest)

ad.test(Ctrposttrial_GAD)
```
A p value greater than .05 indicates the post-trial GAD values for control data are not normally distributed. This is an unexpected result, possibly because the test can be affected by ties in the data. 


```{r}
# Anderson Darling test for normality - Positive Post-trial GAD values for experiment data

#load nortest library
library(nortest)

ad.test(ExpPos_posttrial_GAD)
```
A p value less that .05 indicates positive Post-trial GAD values for experiment data are normally distributed


We decided to proceed with the t-tests as the histograms clearly show the positive data is normally distributed even thought the Anderson Darling test for the control data contradicted this. This requires further investigation possibly by running Shapiro Wilk tests although Anderson Darling is more suited to smaller sample sizes.


```{r}
# Two Sample t-test
t.test(Ctrposttrial_GAD, ExpPos_posttrial_GAD,var.equal = TRUE)

```
The p value is greater than .05 which indicates we should accept the null hypothesis that VR treatment had no effect on GAD scores.

We can say with 95% confidence the true difference in the means between the control and experiment groups is between -1.53 and 0.14.

--- 
Discussion

Limitations

The experiment could have been improved by conducting a double blind experiment where therapists did not know which group the patients were in.

The self-reporting mechanism could be considered an extraneous variable as this type of reporting may yield subjective results.

The fact that there were negative values in the data is concerning especially for a small sample size. It may indicate a lack of care on the part of the inputters or a lack of clarity in the reporting instructions.

Findings

This report finds that the null hypothesis should be accepted as the VR app had no evident effect evident on Generalized Anxiety Disorder 7-item scale scores

---
References



